### 2025-02-21 paper/sections/RelatedWork.tex
add related work section

\section{Related Work}
Multi-label emotion detection has emerged as a significant task in natural language processing (NLP), particularly for low-resource languages. The task is structured in two main output formats: (1) a binary format, which indicates whether an emotion is present in the text, and (2) an intensity scale ranging from 0 to 3, which represents the strength of the emotion in the text.

Given that this task follows a text classification paradigm, various models have been explored to identify the most effective architectures. A considerable amount of research has focused on evaluating different structures to determine the optimal approach. In \cite{wang2016dimensional}, a combination of Long Short-Term Memory (LSTM) networks and Convolutional Neural Networks (CNNs) was explored, where various model configurations were compared based on their F1-score performances. These insights were leveraged to identify suitable model structures for developing a custom model tailored to the specific requirements of this task.

% \section{Related Works}
Several studies have addressed multi-label emotion detection, focusing on various model architectures and preprocessing strategies. In \cite{wang2016dimensional}, a comparative analysis of LSTM and CNN models was performed, demonstrating that hybrid architectures could enhance the F1-score performance of sentiment classification tasks. This research serves as a foundation for designing a model suited to the task at hand.

For low-resource languages, text preprocessing plays a crucial role in improving model performance. The work presented in \cite{muhammad2023semeval} highlighted the effectiveness of multiple preprocessing algorithms specifically designed for African languages. The study demonstrated that well-structured preprocessing pipelines lead to better text representations, ultimately improving classification accuracy.

Moreover, datasets specifically curated for emotion detection in underrepresented languages have been explored. The datasets presented in \cite{muhammad2025brighterbridginggaphumanannotated} and \cite{belay-etal-2025-evaluatin} serve as essential resources for training models and evaluating performance in real-world settings. These datasets enable the training of robust models capable of handling linguistic diversity.

To enhance model performance, modifications to existing architectures have been proposed. Based on the insights from \cite{wang2016dimensional}, additional layers were incorporated into custom models to improve the representation of low-resource languages. This ensures that the models can capture intricate linguistic patterns that might otherwise be overlooked.