### 2025-02-22 paper/tables/MainResults.tex
add results

\begin{table*}[h]
    \centering
    \caption{Results across Track A, B, and C showing macro-average F1 scores of Our Model , Paraticipants Best Model scores, Task Dataset Model\citep{muhammad2025brighterbridginggaphumanannotated} and rankings.}
    \label{tab:results}
    \resizebox{2\columnwidth}{!}{
        \begin{tabular}{l|ccc|cccc|cccc}
            \hline
                              & \multicolumn{3}{c|}{Track A} & \multicolumn{4}{c|}{Track B} & \multicolumn{4}{c}{Track C}                                                                                                                                                     \\
            \cline{2-12}
            \textbf{Language} & \textbf{Our Model}               & \textbf{Best Model}          & \textbf{Rank}               & \textbf{Our Model} & \textbf{Best Model} & \textbf{Task Dataset Model} & \textbf{Rank} & \textbf{Our Model} & \textbf{Best Model} & \textbf{Task Dataset Model} & \textbf{Rank} \\
            \hline
            afr               & 54.01                        & 69.86                        & 15/37                       & —              & —                   & —                & —             & 54.01          & 70.50               & 61.28            & 4/14          \\
            amh               & 61.20                        & 77.31                        & 19/43                       & 49.42          & 85.58               & —                & 14/23         & 61.20          & 66.68               & —                & 6/13          \\
            arq               & 51.07                        & 66.87                        & 24/43                       & 36.54          & 63.38               & 36.37            & 18/26         & 51.07          & 65.35               & 55.75            & 6/14          \\
            ary               & 51.88                        & 62.92                        & 20/41                       & —              & —                   & —                & —             & 51.88          & 63.22               & 52.76            & 6/12          \\
            chn               & 56.65                        & 70.94                        & 30/41                       & 48.47          & 72.24               & 51.86            & 19/27         & 56.65          & 56.65               & 55.23            & 7/15          \\
            deu               & 60.60                        & 73.99                        & 25/50                       & 54.10          & 76.57               & 56.21            & 18/27         & 60.60          & 73.62               & 59.17            & 6/15          \\
            eng               & 73.97                        & 82.30                        & 35/96                       & 68.81          & 84.04               & 64.15            & 28/43         & 73.97          & 82.07               & 65.58            & 4/17          \\
            esp               & 76.19                        & 84.88                        & 26/47                       & 66.70          & 80.80               & 72.59            & 23/29         & 76.19          & 85.00               & 73.29            & 5/15          \\
            hau               & 63.22                        & 75.07                        & 18/40                       & 58.42          & 77.00               & 39.16            & 15/26         & 63.22          & 73.14               & 51.91            & 4/13          \\
            hin               & 80.32                        & 92.57                        & 33/43                       & —              & —                   & —                & —             & 80.32          & 92.16               & 79.73            & 6/16          \\
            ibo               & 50.93                        & 60.01                        & 14/34                       & —              & —                   & —                & —             & 50.93          & 60.47               & 37.40            & 4/12          \\
            ind               & —                            & —                            & —                           & —              & —                   & —                & —             & 35.64          & 67.24               & 57.29            & 15/17         \\
            jav               & —                            & —                            & —                           & —              & —                   & —                & —             & 25.62          & 25.62               & 50.47            & 12/13         \\
            kin               & 51.94                        & 65.74                        & 6/31                        & —              & —                   & —                & —             & 51.94          & 64.59               & 34.36            & 2/11          \\
            mar               & 81.10                        & 90.58                        & 23/42                       & —              & —                   & —                & —             & 81.10          & 90.42               & 77.24            & 6/13          \\
            orm               & 54.31                        & 61.64                        & 11/36                       & —              & —                   & —                & —             & 54.31          & 60.07               & —                & 3/11          \\
            pcm               & 53.09                        & 67.40                        & 22/34                       & —              & —                   & —                & —             & 53.09          & 67.40               & 48.67            & 5/10          \\
            ptbr              & 47.99                        & 68.33                        & 27/42                       & 38.20          & 71.00               & 46.72            & 22/25         & 47.99          & 68.36               & 51.60            & 8/14          \\
            ptmz              & 50.08                        & 54.77                        & 6/36                        & —              & —                   & —                & —             & 50.08          & 55.54               & 40.44            & 3/14          \\
            ron               & 73.75                        & 79.43                        & 14/43                       & 57.61          & 72.60               & 57.69            & 19/26         & 73.75          & 77.27               & 76.23            & 4/15          \\
            rus               & 82.42                        & 90.87                        & 33/50                       & 78.41          & 92.54               & 87.66            & 22/30         & 82.42          & 90.62               & 76.97            & 6/17          \\
            som               & 48.26                        & 57.65                        & 10/34                       & —              & —                   & —                & —             & 48.26          & 56.66               & —                & 4/13          \\
            sun               & 42.48                        & 54.97                        & 19/37                       & —              & —                   & —                & —             & 42.48          & 50.72               & 46.33            & 4/11          \\
            swa               & 29.52                        & 41.47                        & 15/32                       & —              & —                   & —                & —             & 29.52          & 38.43               & 33.27            & 5/14          \\
            swe               & 56.51                        & 62.62                        & 13/40                       & —              & —                   & —                & —             & 56.51          & 64.53               & 51.18            & 6/13          \\
            tat               & 64.32                        & 84.59                        & 18/36                       & —              & —                   & —                & —             & 64.32          & 83.59               & 60.66            & 5/11          \\
            tir               & 52.37                        & 59.05                        & 6/34                        & —              & —                   & —                & —             & 52.37          & 55.24               & —                & 2/10          \\
            ukr               & 48.62                        & 72.56                        & 29/40                       & 42.55          & 70.75               & 43.54            & 17/24         & 48.62          & 71.99               & 54.76            & 10/17         \\
            vmw               & 16.81                        & 32.50                        & 5/31                        & —              & —                   & —                & —             & 16.80          & 26.02               & 20.41            & 6/9           \\
            xho               & —                            & —                            & —                           & —              & —                   & —                & —             & 16.64          & 44.26               & 30.79            & 6/10          \\
            yor               & 34.09                        & 46.13                        & 9/32                        & —              & —                   & —                & —             & 34.09          & 46.79               & 27.44            & 4/11          \\
            zul               & —                            & —                            & —                           & —              & —                   & —                & —             & 16.35          & 16.35               & 22.03            & 8/11          \\
            \hline
        \end{tabular}
    }
\end{table*}