### 2025-02-23 paper/tables/MainResults.tex

% Multilingual-E5-large-instruct \citep{wang2024multilingual}
% AraBert Large Algerian \citep{arabert}