### 2025-02-24 paper/tables/MainResults.tex

\begin{table*}[h]
    \centering
    \caption{Results across Track A, B, and C showing macro-average F1 scores of Our Model , Paraticipants Best Model scores, Task Dataset Model(Baseline)\citep{muhammad2025brighterbridginggaphumanannotated} and rankings.}
    \label{tab:results}
    \resizebox{2\columnwidth}{!}{
        \begin{tabular}{l|c|ccc|cccc|cccc}
            \hline
                                 &                                                                        & \multicolumn{3}{c|}{Track A} & \multicolumn{4}{c|}{Track B} & \multicolumn{4}{c}{Track C}                                                                                                                                         \\
            \cline{3-13}
            \textbf{Language}    & \textbf{Our Model}                                                     & \textbf{Ours}                & \textbf{Best}                & \textbf{Rank}               & \textbf{Ours} & \textbf{Best} & \textbf{Baseline} & \textbf{Rank} & \textbf{Ours} & \textbf{Best} & \textbf{Baseline} & \textbf{Rank} \\
            \hline
            Afrikaans(afr)       & \citep{wang2024multilingual} + SVM                                     & 54.01                        & 69.86                        & 15/37                       & —             & —             & —                 & —             & 54.01         & 70.50         & 61.28             & 4/14          \\
            Amharic(amh)         & \citep{arabert}               + SVM                                    & 61.20                        & 77.31                        & 19/43                       & 49.42         & 85.58         & —                 & 14/23         & 61.20         & 66.68         & —                 & 6/13          \\
            Algerian Arabic(arq) & \citep{wang2024multilingual}   + SVM                                   & 51.07                        & 66.87                        & 24/43                       & 36.54         & 63.38         & 36.37             & 18/26         & 51.07         & 65.35         & 55.75             & 6/14          \\
            Moroccan Arabic(ary) & \citep{wang2024multilingual}   + SVM                                   & 51.88                        & 62.92                        & 20/41                       & —             & —             & —                 & —             & 51.88         & 63.22         & 52.76             & 6/12          \\
            Chinese(chn)         & \citep{zpoint-large-embedding-zh} + SVM                                & 56.65                        & 70.94                        & 30/41                       & 48.47         & 72.24         & 51.86             & 19/27         & 56.65         & 56.65         & 55.23             & 7/15          \\
            German(deu)          & \citep{wang2024multilingual}       + SVM                               & 60.60                        & 73.99                        & 25/50                       & 54.10         & 76.57         & 56.21             & 18/27         & 60.60         & 73.62         & 59.17             & 6/15          \\
            English(eng)         & \citep{zhang2025jasperstelladistillationsota}  + SVM                   & 73.97                        & 82.30                        & 35/96                       & 68.81         & 84.04         & 64.15             & 28/43         & 73.97         & 82.07         & 65.58             & 4/17          \\
            Spanish(esp)         & \citep{wang2024multilingual}      + SVM                                & 76.19                        & 84.88                        & 26/47                       & 66.70         & 80.80         & 72.59             & 23/29         & 76.19         & 85.00         & 73.29             & 5/15          \\
            Hausa(hau)           & \citep{dobler-demelo-2023-focus} + SVM                                 & 63.22                        & 75.07                        & 18/40                       & 58.42         & 77.00         & 39.16             & 15/26         & 63.22         & 73.14         & 51.91             & 4/13          \\
            Hindi(hin)           & \citep{wang2024multilingual}  + SVM                                    & 80.32                        & 92.57                        & 33/43                       & —             & —             & —                 & —             & 80.32         & 92.16         & 79.73             & 6/16          \\
            Igbo(ibo)            & \citep{wang2024multilingual}   + SVM                                   & 50.93                        & 60.01                        & 14/34                       & —             & —             & —                 & —             & 50.93         & 60.47         & 37.40             & 4/12          \\
            Indonesian(ind)      & \citep{wang2024multilingual}   + XGB                                   & —                            & —                            & —                           & —             & —             & —                 & —             & 35.64         & 67.24         & 57.29             & 15/17         \\
            Javanese(jav)        & \citep{wang2024multilingual}   + XGB                                   & —                            & —                            & —                           & —             & —             & —                 & —             & 25.62         & 25.62         & 50.47             & 12/13         \\
            Kinyarwanda(kin)     & \citep{wang2024multilingual}  + SVM                                    & 51.94                        & 65.74                        & 6/31                        & —             & —             & —                 & —             & 51.94         & 64.59         & 34.36             & 2/11          \\
            Marathi(mar)         & \citep{wang2024multilingual}   + SVM                                   & 81.10                        & 90.58                        & 23/42                       & —             & —             & —                 & —             & 81.10         & 90.42         & 77.24             & 6/13          \\
            Oromo(orm)           & \citep{wang2024multilingual}    + SVM                                  & 54.31                        & 61.64                        & 11/36                       & —             & —             & —                 & —             & 54.31         & 60.07         & —                 & 3/11          \\
            Nigerian-Pidgin(pcm) & \citep{wang2024multilingual}   + SVM                                   & 53.09                        & 67.40                        & 22/34                       & —             & —             & —                 & —             & 53.09         & 67.40         & 48.67             & 5/10          \\
            Pt* Brazilian(ptbr)  & \citep{souza2020bertimbau} + SVM                                       & 47.99                        & 68.33                        & 27/42                       & 38.20         & 71.00         & 46.72             & 22/25         & 47.99         & 68.36         & 51.60             & 8/14          \\
            Pt* Mozambican(ptmz) & \citep{wang2024multilingual}   + SVM                                   & 50.08                        & 54.77                        & 6/36                        & —             & —             & —                 & —             & 50.08         & 55.54         & 40.44             & 3/14          \\
            Romanian(ron)        & \citep{wang2024multilingual}     + SVM                                 & 73.75                        & 79.43                        & 14/43                       & 57.61         & 72.60         & 57.69             & 19/26         & 73.75         & 77.27         & 76.23             & 4/15          \\
            Russian(rus)         & \citep{snegirev2025russianfocusedembeddersexplorationrumteb}     + SVM & 82.42                        & 90.87                        & 33/50                       & 78.41         & 92.54         & 87.66             & 22/30         & 82.42         & 90.62         & 76.97             & 6/17          \\
            Somali(som)          & \citep{wang2024multilingual}       + SVM                               & 48.26                        & 57.65                        & 10/34                       & —             & —             & —                 & —             & 48.26         & 56.66         & —                 & 4/13          \\
            Sundanese(sun)       & \citep{wang2024multilingual}      + SVM                                & 42.48                        & 54.97                        & 19/37                       & —             & —             & —                 & —             & 42.48         & 50.72         & 46.33             & 4/11          \\
            Swahili(swa)         & \citep{wang2024multilingual}      + SVM                                & 29.52                        & 41.47                        & 15/32                       & —             & —             & —                 & —             & 29.52         & 38.43         & 33.27             & 5/14          \\
            Swedish(swe)         & \citep{wang2024multilingual}     + SVM                                 & 56.51                        & 62.62                        & 13/40                       & —             & —             & —                 & —             & 56.51         & 64.53         & 51.18             & 6/13          \\
            Tatar(tat)           & \citep{wang2024multilingual}    + SVM                                  & 64.32                        & 84.59                        & 18/36                       & —             & —             & —                 & —             & 64.32         & 83.59         & 60.66             & 5/11          \\
            Tigrinya(tir)        & \citep{wang2024multilingual}    + SVM                                  & 52.37                        & 59.05                        & 6/34                        & —             & —             & —                 & —             & 52.37         & 55.24         & —                 & 2/10          \\
            Ukrainian(ukr)       & \citep{sturua2024jinaembeddingsv3multilingualembeddingstask}   + SVM   & 48.62                        & 72.56                        & 29/40                       & 42.55         & 70.75         & 43.54             & 17/24         & 48.62         & 71.99         & 54.76             & 10/17         \\
            Emakhuwa(vmw)        & \citep{sturua2024jinaembeddingsv3multilingualembeddingstask}   + SVM   & 16.81                        & 32.50                        & 5/31                        & —             & —             & —                 & —             & 16.80         & 26.02         & 20.41             & 6/9           \\
            isiXhosa(xho)        & \citep{wang2024multilingual}     + XGB                                 & —                            & —                            & —                           & —             & —             & —                 & —             & 16.64         & 44.26         & 30.79             & 6/10          \\
            Yoruba(yor)          & \citep{wang2024multilingual}     + SVM                                 & 34.09                        & 46.13                        & 9/32                        & —             & —             & —                 & —             & 34.09         & 46.79         & 27.44             & 4/11          \\
            isiZulu(zul)         & \citep{wang2024multilingual}     + XGB                                 & —                            & —                            & —                           & —             & —             & —                 & —             & 16.35         & 16.35         & 22.03             & 8/11          \\
            \hline
            % \\
            % \bottomrule
        \end{tabular}
    }
    \\
    \footnotesize{\textit{*: pt=Portuguese}}
\end{table*}