### 2025-02-24 paper/appendix/ModelSelection.tex
add appendix

\section{Model Selection}
\label{sec:appendix_model_selection}
The training dataset of Track A was utilized for model selection, where it was split into an 80/20 ratio to create a training set and a validation set. Several models were evaluated on this split to identify the optimal model based on performance metrics such as recall, precision, and F1-score. The model that achieved the best balance across these metrics was selected as the final model and applied consistently across all three tracks—A, B, and C. A detailed comparative analysis of the performance of different models for each language is presented in Table \ref{tab:model_selection}, where each row corresponds to a specific language and includes results for multiple models, with columns reporting recall, precision, and F1-score for each.

\input{tables/ModelSelection}

Table \ref{tab:model_selection} provides a comprehensive comparison of the models' performance for each language, aiding in the selection of the final model applied to Tracks A, B, and C.