### 2025-02-27 paper/tables/ErrorAnalysis.tex

\begin{table*}[h!]
    \centering
    \resizebox{2\columnwidth}{!}{
        \begin{tabular}{|p{12cm}|c|c|c|c|c|c|}
            \hline
            \textbf{Text}                                                                                                                                                                                                                                                                           & \textbf{Type} & \textbf{Anger} & \textbf{Fear} & \textbf{Joy} & \textbf{Sadness} & \textbf{Surprise} \\ \hline
            \multirow{2}{12cm}{I'm just numb.}                                                                                                                                                                                                                                                      & Truth         & 0              & 0             & 0            & 1                & 0                 \\ \cline{2-7}
                                                                                                                                                                                                                                                                                                    & Pred          & 0              & 1             & 0            & 1                & 0                 \\ \hline
            \multirow{2}{12cm}{At the time it didn't seem to bother me.}                                                                                                                                                                                                                            & Truth         & 0              & 0             & 0            & 1                & 0                 \\ \cline{2-7}
                                                                                                                                                                                                                                                                                                    & Pred          & 0              & 0             & 0            & 0                & 0                 \\ \hline
            \multirow{2}{12cm}{I found out six weeks before                                                                                                                     the wedding that my dad had only six weeks to live (he had cancer for two years... a fact she was fully aware of).} & Truth         & 1              & 1             & 0            & 1                & 1                 \\ \cline{2-7}
                                                                                                                                                                                                                                                                                                    & Pred          & 0              & 1             & 0            & 1                & 0                 \\ \hline
        \end{tabular}
    }
    \caption{Error Analysis Table of language English for track A}
    \label{tab:error_analysis}
\end{table*}