### 2025-02-27 paper/tables/MainResults.tex
update rankings

\begin{table*}[h]
    \centering
    \caption{Results across Track A, B, and C showing macro-average F1 scores of Our Model , Paraticipants Best Model scores, Task Dataset Best Model with Baseline\citep{muhammad2025brighterbridginggaphumanannotated} and rankings.}
    \label{tab:results}
    \resizebox{2\columnwidth}{!}{
        \begin{tabular}{l|c|cccc|ccccc|ccccc}
            \hline
                                   &                                                                        & \multicolumn{4}{c|}{Track A} & \multicolumn{5}{c|}{Track B} & \multicolumn{5}{c}{Track C}                                                                                                                                                                               \\
            \cline{3-16}
            \textbf{Language}      & \textbf{Our Model}                                                     & \textbf{Ours}                & \textbf{BP*}                 & \textbf{Base}               & \textbf{Rank} & \textbf{Ours} & \textbf{BP*} & \textbf{BDP**} & \textbf{Base} & \textbf{Rank} & \textbf{Ours} & \textbf{BP*} & \textbf{BDP**} & \textbf{Base} & \textbf{Rank} \\
            \hline
            Afrikaans(afr)         & \citep{wang2024multilingual} + SVM                                     & 54.01                        & 69.86                        & 37.14                       & 15/39         & —             & —           & —              & —             & —             & 54.01         & 70.50       & 61.28          & 35.04         & 4/15          \\
            Amharic(amh)           & \citep{arabert}               + SVM                                    & 61.20                        & 77.31                        & 63.83                       & 20/45         & 49.42         & 85.58       & —              & 50.79         & 15/24         & 61.20         & 66.68       & —              & 48.66         & 6/14          \\
            Algerian Arabic(arq)   & \citep{wang2024multilingual}   + SVM                                   & 51.07                        & 66.87                        & 41.41                       & 24/45         & 36.54         & 64.97       & 36.37          & 1.64          & 18/27         & 51.07         & 65.35       & 55.75          & 33.78         & 6/15          \\
            Moroccan Arabic(ary)   & \citep{wang2024multilingual}   + SVM                                   & 51.88                        & 62.92                        & 47.16                       & 20/43         & —             & —           & —              & —             & —             & 51.88         & 63.22       & 52.76          & 35.46         & 6/13          \\
            Chinese(chn)           & \citep{zpoint-large-embedding-zh} + SVM                                & 56.65                        & 70.94                        & 53.08                       & 30/43         & 48.47         & 72.24       & 51.86          & 40.53         & 19/28         & 56.65         & 68.89       & 55.23          & 24.56         & 7/16          \\
            German(deu)            & \citep{wang2024multilingual}       + SVM                               & 60.60                        & 73.99                        & 64.23                       & 26/52         & 54.10         & 76.57       & 56.21          & 56.21         & 19/28         & 60.60         & 73.62       & 59.17          & 46.84         & 6/16          \\
            English(eng)           & \citep{zhang2025jasperstelladistillationsota}  + SVM                   & 73.97                        & 82.30                        & 70.83                       & 35/98         & 68.81         & 84.04       & 64.15          & 64.15         & 28/45         & 73.97         & 82.07       & 65.58          & 37.54         & 4/18          \\
            Spanish(esp)           & \citep{wang2024multilingual}      + SVM                                & 76.19                        & 84.88                        & 77.44                       & 28/49         & 66.70         & 80.80       & 72.59          & 72.59         & 24/30         & 76.19         & 85.00       & 73.29          & 57.37         & 5/16          \\
            Hausa(hau)             & \citep{dobler-demelo-2023-focus} + SVM                                 & 63.22                        & 75.07                        & 59.55                       & 18/42         & 58.42         & 77.00       & 39.16          & 27.03         & 15/27         & 63.22         & 73.14       & 51.91          & 31.98         & 4/14          \\
            Hindi(hin)             & \citep{wang2024multilingual}  + SVM                                    & 80.32                        & 92.57                        & 85.51                       & 35/45         & —             & —           & —              & —             & —             & 80.32         & 92.16       & 79.73          & 13.75         & 6/17          \\
            Igbo(ibo)              & \citep{wang2024multilingual}   + SVM                                   & 50.93                        & 60.01                        & 47.90                       & 14/36         & —             & —           & —              & —             & —             & 50.93         & 60.47       & 37.40          & 50.93         & 4/13          \\
            Indonesian(ind)        & \citep{wang2024multilingual}   + XGB                                   & —                            & —                            & —                           & —             & —             & —           & —              & —             & —             & 35.64         & 67.24       & 57.29          & 37.64         & 16/18         \\
            Javanese(jav)          & \citep{wang2024multilingual}   + XGB                                   & —                            & —                            & —                           & —             & —             & —           & —              & —             & —             & 25.62         & 46.38       & 50.47          & 46.38         & 13/14         \\
            Kinyarwanda(kin)       & \citep{wang2024multilingual}  + SVM                                    & 51.94                        & 65.74                        & 46.29                       & 6/33          & —             & —           & —              & —             & —             & 51.94         & 64.59       & 34.36          & 18.38         & 2/12          \\
            Marathi(mar)           & \citep{wang2024multilingual}   + SVM                                   & 81.10                        & 90.58                        & 82.20                       & 25/44         & —             & —           & —              & —             & —             & 81.10         & 90.42       & 77.24          & 77.24         & 6/14          \\
            Oromo(orm)             & \citep{wang2024multilingual}    + SVM                                  & 54.31                        & 61.64                        & 12.63                       & 11/38         & —             & —           & —              & —             & —             & 54.31         & 60.07       & —              & 26.17         & 3/12          \\
            Nigerian-Pidgin(pcm)   & \citep{wang2024multilingual}   + SVM                                   & 53.09                        & 67.40                        & 55.50                       & 23/36         & —             & —           & —              & —             & —             & 53.09         & 67.40       & 48.67          & 1.01          & 5/11          \\
            Pt*** Brazilian(ptbr)  & \citep{souza2020bertimbau} + SVM                                       & 47.99                        & 68.33                        & 42.57                       & 28/44         & 38.20         & 71.00       & 46.72          & 29.74         & 22/26         & 47.99         & 68.36       & 51.60          & 41.84         & 8/15          \\
            Pt*** Mozambican(ptmz) & \citep{wang2024multilingual}   + SVM                                   & 50.08                        & 54.77                        & 45.91                       & 6/38          & —             & —           & —              & —             & —             & 50.08         & 55.54       & 40.44          & 29.67         & 3/15          \\
            Romanian(ron)          & \citep{wang2024multilingual}     + SVM                                 & 73.75                        & 79.43                        & 76.23                       & 15/45         & 57.61         & 72.60       & 57.69          & 55.66         & 18/27         & 73.75         & 77.27       & 76.23          & 76.23         & 5/16          \\
            Russian(rus)           & \citep{snegirev2025russianfocusedembeddersexplorationrumteb}     + SVM & 82.42                        & 90.87                        & 83.77                       & 34/52         & 78.41         & 92.54       & 87.66          & 87.66         & 23/31         & 82.42         & 90.62       & 76.97          & 70.43         & 6/18          \\
            Somali(som)            & \citep{wang2024multilingual}       + SVM                               & 48.26                        & 57.65                        & 45.93                       & 10/36         & —             & —           & —              & —             & —             & 48.26         & 56.66       & —              & 27.27         & 4/14          \\
            Sundanese(sun)         & \citep{wang2024multilingual}      + SVM                                & 42.48                        & 54.97                        & 37.31                       & 20/39         & —             & —           & —              & —             & —             & 42.48         & 50.72       & 46.33          & 19.43         & 4/12          \\
            Swahili(swa)           & \citep{wang2024multilingual}      + SVM                                & 29.52                        & 41.47                        & 22.65                       & 16/34         & —             & —           & —              & —             & —             & 29.52         & 38.43       & 33.27          & 18.99         & 5/15          \\
            Swedish(swe)           & \citep{wang2024multilingual}     + SVM                                 & 56.51                        & 62.62                        & 51.98                       & 13/42         & —             & —           & —              & —             & —             & 56.51         & 64.53       & 51.18          & 51.18         & 6/14          \\
            Tatar(tat)             & \citep{wang2024multilingual}    + SVM                                  & 64.32                        & 84.59                        & 53.94                       & 19/38         & —             & —           & —              & —             & —             & 64.32         & 83.59       & 60.66          & 44.54         & 5/12          \\
            Tigrinya(tir)          & \citep{wang2024multilingual}    + SVM                                  & 52.37                        & 59.05                        & 46.28                       & 6/36          & —             & —           & —              & —             & —             & 52.37         & 55.24       & —              & 33.93         & 2/11          \\
            Ukrainian(ukr)         & \citep{sturua2024jinaembeddingsv3multilingualembeddingstask}   + SVM   & 48.62                        & 72.56                        & 53.45                       & 31/42         & 42.55         & 70.75       & 43.54          & 39.94         & 17/25         & 48.62         & 71.99       & 54.76          & 49.56         & 11/18         \\
            Emakhuwa(vmw)          & \citep{sturua2024jinaembeddingsv3multilingualembeddingstask}   + SVM   & 16.81                        & 32.50                        & 12.14                       & 15/33         & —             & —           & —              & —             & —             & 16.80         & 26.02       & 20.41          & 5.22          & 6/10          \\
            isiXhosa(xho)          & \citep{wang2024multilingual}     + XGB                                 & —                            & —                            & —                           & —             & —             & —           & —              & —             & —             & 16.64         & 44.26       & 30.79          & 12.73         & 6/11          \\
            Yoruba(yor)            & \citep{wang2024multilingual}     + SVM                                 & 34.09                        & 46.13                        & 9.22                        & 9/34          & —             & —           & —              & —             & —             & 34.09         & 46.79       & 27.44          & 5.33          & 4/12          \\
            isiZulu(zul)           & \citep{wang2024multilingual}     + XGB                                 & —                            & —                            & —                           & —             & —             & —           & —              & —             & —             & 16.35         & 39.69       & 22.03          & 15.26         & 8/12          \\
            \hline
            % \\
            % \bottomrule
        \end{tabular}
    }
    \\
    \footnotesize{\textit{BP*=result of rank 1\\BDP**=best result of dataset paper\citep{muhammad2025brighterbridginggaphumanannotated}\\pt***=Portuguese}}
    \\
    \footnotesize{The Best result of dataset paper for Track A is identical to that of Track C.}
\end{table*}